---
title: "DATA 607 Project 1"
author: "Leslie Tavarez"
date: "2024-10-06"
output: html_document
---
##### Introduction: In this project, we are given a text file with chess tournament results. This text file has a specific structure. Our goal is to create an R Markdown file that generates a .CSV file with the following info for all of the players: "Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents". First, I will load all necessary libraries and attempt to clean the data as much as I can. 



```{r}
library(dplyr)
library(readr)
library(stringr)


# Read the data without headers
raw_data <- read.table("/Users/leslie/Downloads/tournamentinfo.txt", 
                       sep = "|", 
                       fill = TRUE, 
                       stringsAsFactors = FALSE)

# Check the structure of the raw data
str(raw_data)

# Initial column assignment
colnames(raw_data) <- c("Pair_Num", "Player_Name", "Total", "Round_1", "Round_2", "Round_3", "Round_4", "Round_5", "Round_6", "Round_7")

# Remove the first four rows using indexing
raw_data <- raw_data[-c(1:4), ]

# Remove the last column if it contains only NA values
if (all(is.na(raw_data[, ncol(raw_data)]))) {
  raw_data <- raw_data[, -ncol(raw_data)]
}

# Remove rows that contain only dashes
raw_data <- raw_data[!grepl("^\\s*-{2,}\\s*$", raw_data$Pair_Num), ]

View(raw_data)
```
##### Now I am going to calulate the average opponent rating, extract necessary information like player name, player state, total points and pre-rating and save into csv file. 
```{r}
# Function to calculate average opponent rating
calculate_average_opponent_rating <- function(rounds) {
  opponent_ratings <- as.numeric(unlist(str_extract_all(rounds, "\\d+")))
  mean(opponent_ratings, na.rm = TRUE)
}

# Extract Player's State and Pre-Rating from Player_Name
extract_player_info <- function(name) {
  parts <- str_split(name, "\\s+/\\s+")[[1]]
  state <- str_trim(parts[1])
  rating_info <- str_extract(parts[2], "\\d+ -> \\d+")
  pre_rating <- as.numeric(str_extract(rating_info, "\\d+"))
  return(c(state, pre_rating))
}

# Create an empty data frame for results
results <- data.frame(
  Player_Name = character(),
  Player_State = character(),
  Total_Points = numeric(),
  Pre_Rating = numeric(),
  Average_Opponent_Rating = numeric(),
  stringsAsFactors = FALSE
)
# Clean the Total column
raw_data$Total <- gsub("\\s+", "", raw_data$Total)  # Remove extra whitespace
raw_data$Total <- gsub("[^0-9.]", "", raw_data$Total)  # Remove non-numeric characters

# Convert to numeric, suppress warnings
raw_data$Total <- suppressWarnings(as.numeric(raw_data$Total))

sum(is.na(raw_data$Total))



# Loop through each player
for (i in 1:nrow(raw_data)) {
  player_name <- raw_data$Player_Name[i]

# Use the cleaned Total value
  total_points <- raw_data$Total[i]
  
# Extract Player's State from Pair_Num (two capital letters)
  player_state <- str_extract(raw_data$Pair_Num[i], "[A-Z]{2}")
  
# Extract Pre-Rating from Player_Name
  pre_rating_match <- str_extract(player_name, "R:\\s*(\\d+)")
  pre_rating <- as.numeric(str_extract(pre_rating_match, "\\d+"))
  
# Get total points (ensure it's numeric)
  total_points <- as.numeric(raw_data$Total[i])
  
# Calculate average opponent rating
  rounds <- unlist(raw_data[i, 4:ncol(raw_data)])
  avg_opponent_rating <- calculate_average_opponent_rating(rounds[!is.na(rounds)])
 

  
  # Add to results data frame
  results <- rbind(results, data.frame(
    Player_Name = player_name,
    Player_State = player_state,
    Total_Points = total_points,
    Pre_Rating = pre_rating,
    Average_Opponent_Rating = avg_opponent_rating
  ))
}
 

View(results)

# Save the results to a CSV file
write.csv(results, "chess_tournament_results.csv", row.names = FALSE)
```

##### Conclusion: I was able to load the text file, clean it up a little, and create a csv file. But now I have this problem were I still have data I do not need in my results. Ive tried to remove the USCF ID/ Rtg (Pre->Post) but whenever I do, it messes up my column with the pre-rating. This is leading to extra rows that I do not need with "NA" values. I cannot remove entire row with "NA" or "NaN" because it still has important information that I cannot remove from my data table. I've tried many different things and nothing has worked. I am not sure what else to do here. 
